\documentclass[a4paper,12pt,times,print,index,custombib,custommargin]{PhDThesisPSnPDF}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
%\documentclass[a4paper,12pt,times,numbered,print,index]{PhDThesisPSnPDF}

\usepackage{import}

\input{../Preamble/preamble}

%******************************** Glossary *********************************************
\input{../Glossary/glossary}

%%%%%% -- KNITR SETUP -- %%%%%%%%%%

%************************************************************************************************
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\tableofcontents

\listoffigures

\listof{cmd}{List of commands}

%\todototoc
\listoftodos

\printglossaries


% **************************** Define Graphics Path **************************
%% Note, every path needs to end with a "/" !!!
\ifpdf
    \graphicspath{
    {./Figs/Raster/}
    {./Figs/PDF/}
    {./Figs/}
    {../Data_analysis/reference-mapping/figure/}
    }
\else
    \graphicspath{ 
    {./Figs/Vector/}
    {./Figs/}
%    {/Users/Claudius/Documents/PhD/THESIS/kks32/LaTeX/3_Chapter/Figs/}
    }
\fi

%
%
%
\chapter{Investigation into the demographic history of the hybrid zone}
%
%
%


\abstract{
In this chapter I want to use all RAD data from all different libraries that I created to investigate the evolutionary history of the hybrid zone.

I will first need to assemble the RAD reads of the different libraries. For that, I want to look into and test a few different programmes, pipelines and analysis protocols that have recently been published. I have created a list of 14 priority papers in my literature library in JabRef under the category \textsf{sRAD} $\rightarrow$ \textsf{RAD de novo assembly}.

I will need to estimate allele frequencies from the RAD data for the two populations I got data for and use it in an ABC analysis. Parameters of interest would be:

\begin{itemize}
\item time of separation
\item recent migration rate / rate of gene flow 
\item recent bottlenecks
\item signals of selection
\end{itemize}

I will need to create a list of papers to read about the technical details of ABC analysis. For the implementation, I will need the assistance of Ludovic.

Finally, since I am trying to extend the scarce knowledge about the history of the hybrid zone, I will need to write a little literature review about what is known so far. Most of these papers should be there as printed copies in the hybrid zone paper box. 

%\newcommand{\Todo}[1]{\todo[size=\normalsize,nolist,inline]{#1}} % don't know why \renewcommand doesn't work
\newcommand{\Todo}[2][orange]{\todo[size=\normalsize,nolist,inline,color=#1]{#2}} % don't know why \renewcommand doesn't work

\begin{itemize}
\item RAD de novo assembly:
	\begin{itemize}
	\item \Todo[green]{create list of papers to read}
	\item \Todo[green]{read through list of papers and select programmes to test}
	\item \Todo{test selected programmes and pipelines}
	\item \Todo{Get an estimate of the sequencing error rate from the standard RAD data.}
	\item \Todo{run \texttt{RepeatMasker} on Radome}
	\item \Todo{screen Radome for \textit{Wolbachia} sequences, maybe by doing a proper PE read assembly again with a proper blast annotation}
	\item \Todo{assemble different libraries}
	\item \Todo{estimate allele frequencies of SNP's and indels in a Bayesian framework}
	\end{itemize}
\item ABC:
	\begin{itemize}
	\item \Todo{explore \texttt{ms}}
	\item \Todo{create list of papers to read}
	\item \Todo{test selected programmes and pipelines}
	\item \Todo{account for biases in the data: null alleles}	
	\item \Todo{design ABC analysis: parameters}
	\item \Todo{familiarise yourself with Ludovic's ABC pipeline on Iceberg}
	\item \Todo{do ABC analysis}
	\item \Todo{analyse posterior distributions of parameters}
	\end{itemize}
\item Literature review of hybrid zone history:
	\begin{itemize}
	\item \Todo{create list of papers to read}
	\item \Todo{read and digest papers while writing the review}
	\end{itemize}
\end{itemize}
}

%----------------------------
% Bibliography
%----------------------------

\bibliographystyle{elsarticle-harv}
\bibliography{/Users/Claudius/Documents/MyLiterature/Literature}

\end{document}

