---
title: "Sequence divergene from 2D SFS"
author: "Claudius"
date: "16/01/2018"
output: 
  #html_document
  github_document
bibliography: Lit.bib
---

```{r setup, cache=FALSE, include=TRUE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE, dev=c("png", "pdf"), eval=TRUE, fig.width=10, fig.height=8, warning = FALSE)
options(digits=10)
setwd("/Users/Claudius/Documents/PhD/THESIS/kks32/LaTeX/Data_analysis/SNP-indel-calling/ANGSD/BOOTSTRAP_CONTIGS/minInd9_overlapping/SFS")
```

```{r, echo=FALSE, include=FALSE}
library(fields)
```

```{r}
# plot 2D SFS
sfs2d = scan("original/EryPar.unfolded.sfs")
sfs2d = t(matrix(sfs2d, nrow=37, ncol=37)) # rows should be PAR, columns should be ERY
sfs2d[1,1] = 0
ticks = c(1, 10, 100, 500, 1000, 5000, 10000)
# rows in the matrix are on the x-axis, columns are on the y-axis:
image.plot(0:37, 0:37, log10(sfs2d+1), xlab="non-reference sample allele frequency in PAR", 
           ylab="non-reference sample allele frequency in ERY", main="global unfolded 2D-SFS",
           axis.args = list(at=log10(ticks), labels=ticks), legend.lab="number of SNP's",
           legend.line = 3,
           graphics.reset=TRUE
           )
```


I would like to calculate $K_{ST}$ as:
$$
K_{ST} = 1 - \frac{K_S}{K_T}
$$

$K_S$ is the average within population $\pi$ from ERY and PAR, weighted by sample size:
$$
K_S = \frac{n}{n+m} \times \pi_{ery} + \frac{m}{n+m} \times \pi_{par}
$$

where $n$ is the sample size of ERY (36 alleles) and $m$ is the sample size of PAR (36 alleles). $\pi_{ery}$ and $\pi_{par}$ can be calculated from the 1D spectra taken from the 2D SFS by marginalising over one dimension, i. e. `colSums` or `rowSums`. $K_T$ is:
$$
K_T = K_S + K_{B}
$$
$K_B$ is the average pairwise sequence divergence between ERY and PAR. It can be calculated from the unfolded 2D SFS with:
$$
K_B = \frac{1}{nm} \left[ \sum_{i=0}^{m} \sum_{j=0}^{n} [i(n-j) + j(m-i)]\xi_{ij} \right]
$$
where $\xi_{ij}$ is the count in the unfolded joint allele frequency class [i, j] in the 2D-SFS. The allele frequency classes in the 2D SFS refer to the non-reference allele. The joint reference allele frequency is given by [m-i, n-j]. $K_B$ should be equivalent to $d_{xy}$ as given in eq. 12.66 on p. 256 in Nei & Kumar (2000).

```{r}
get_Kb = function(sfs2d){
  m = dim(sfs2d)[1]-1
  n = dim(sfs2d)[2]-1
  Kb = 0
  for(i in 0:n){
    for(j in 0:m){
      Kb = Kb + (i*(n-j)+j*(m-i))*sfs2d[i+1,j+1]
    }
  }
  return(Kb/(n*m))
}
```


```{r}
# get 2D SFS including count for [0, 0]
sfs2d = scan("original/EryPar.unfolded.sfs")
sfs2d = t(matrix(sfs2d, nrow=37, ncol=37)) # rows should be PAR, columns should be ERY
# get number of sites over which 2D SFS has been estimated
n.Sites = sum(sfs2d)
```


```{r}
Kb = get_Kb(sfs2d)/n.Sites
Kb
```
The average sequence divergence between ERY and PAR is 10.7%. This average sequence divergence includes mutations that are not fixed in either or both subspecies.


```{r, cache=TRUE}
# define function to calculate average number of pairwise differences
# from a unfolded SFS

PI_unfolded = function(sfs){
  # takes UN-folded SFS
  # expects that first count is for reference alleles
  n = length(sfs)-1
  # modified from equation 1.4 in Wakeley2009 for unfolded SFS:
  2/(n*(n-1)) * sum( sapply(1:n, function(i) i*(n-i)*sfs[i+1]) )
}
```

```{r}
# marginalise
# rowSums should give the marginal spectrum for PAR, colSums should give the marginal spectrum for ERY

marginalise = function(sfs2d, over){
  # if 'over' is 1, returns the marginal spectrum for ERY,
  # if 'over' is 2, returns the marginal spectrum for PAR
  if(over==2){
    return(rowSums(sfs2d))
  }else{
    if(over==1){
      return(colSums(sfs2d))
    }else{
      print("over must be given 1 or 2")
      }
  }
}
```


```{r}
# get within population pi for PAR
PI_unfolded(marginalise(sfs2d, over=2))/n.Sites
# get within population pi for ERY
PI_unfolded(marginalise(sfs2d, over=1))/n.Sites
```
These values are very similar to those estimated from 1D spectra of overlapping sites (see table 1.1 in thesis).

```{r}
n = 36
m = 36
Ks = 36/72 * PI_unfolded(marginalise(sfs2d, over=2))/n.Sites + 36/72 * PI_unfolded(marginalise(sfs2d, over=1))/n.Sites
Ks
```

```{r}
Kt = Ks + Kb
Kt
```

```{r}
Kst = 1 - Ks/Kt
Kst
```

This $K_{ST}$ estimate is roughly concordant with estimates of $K_{ST}$ between French and Spanish regions given in Cooper (1995). 


The net sequence divergence between ERY and PAR is:
$$
D_a = K_B - K_S
$$

see equation 12.67, p. 256 in Nei & Kumar (2000).

```{r}
Da = Kb - Ks
Da
```

$$
E[D_a] = 2\mu T
$$
eq. 12.69, p. 256 in Nei & Kumar (2000).

```{r}
T = Da/2/3*10^9
T
```
The divergence time as estimated from genome-wide $D_a$ is close to 500,000 years (assuming a mutation rate of $3\times 10^{-9}$). This is remarkably close to the estimate given by dadi in the _divergence-in-isolation_ model (486,848, CI:476,703 – 496,994). 


```{r}
(sum(sfs2d[37,1:37]) + sum(sfs2d[1:37,37]))/n.Sites
```

